% Place acknowledgments below.
I would like to express my most profound gratitude to my doctoral advisor Dr. Rahul Rai, for his unlimited guidance, care, support and constant vigil throughout my Ph.D. studies at the University at Buffalo. I am grateful to him for introducing me to the field of Uncertainty Quantification. I wish to thank my committee and doctoral co-advisors. My weekly meetings with them in this four and half years helped me to complete my dissertation fast.

I wish to thank Dr. Puneet Singla for his insights on doing high-quality research on dynamical systems and stochastic filtering. He has taught me everything required for this work and helped me through developing the main backbone of this study. I would like to express my gratitude to Dr. Tarunraj Singh for his detailed criticism since the very beginning of my Ph.D. and helped me get a better insight into the requirements of this research. He provided me with the shallow-water equation model
and provided constant feedback while working on it. I thank Dr. Abani Patra for his help with my research and his knowledge of functional analysis and his constructive input while working on the geophysical mass flow problem. It is a pleasure working with all four professors. All the weekly meetings that I had in throughout my Ph.D. helped me get a good shape of this dissertation.

I thank my collaborator Anna Schweda for helping me to create the building energy model that I have used as an application problem in my thesis. My gratitude also extends to Dr. Sonjoy Das for his course on Probability and Random Matrix. I learned a lot during these courses, which helped me to complete my dissertation problem.

I thank my childhood friends Debashish, Riya, and Ishani for standing by me in my good and bad times. 
My life in Buffalo has been fascinating. Since the very onset, I have been a part of a cultural community and have spent a lot of time in doing cultural activities (often more than required). I met some excellent people here out of this University who helped me a lot in my hour of needs. I want to thank Mriganka, Debrup, Dithi, Srirupa, Anudipa, Debanjan, Navneet, Kundan, Supratik, Arinjoy, Umaer, Ranjit, and Sarthak. I sincerely apologize if I have missed out any one. 

Throughout my life, I have met several mathematics enthusiasts. I would like to start with my maternal-uncle Mr. Tapas Roy, who is a high-school mathematics teacher by profession. He helped me to learn a lot of mathematics in my early school life and guided me in learning preliminary computer programming as well. I extend my gratitude to my high-school tutor Mr. Hafizul Mondal who instilled in me the love of mathematics. I was immensely inspired by another tutor Dr. Arun Sanyal, who changed my perception towards the subject and who taught me new ways to tackle any challenging problem. I thank my Masters' thesis supervisor from the Indian Statistical Institute Dr. Sanghamitra Bandyopadhyay for my first lesson in research. Her words of wisdom and advice have been of great help in all my research works. I also thank Dr. Malay Bhattacharyya for his countless advice.    

I thank my parents for their constant support throughout this time. They were my biggest inspiration. They have worked hard in my early academic life, during my school time. They pushed me to my limits and made me challenge myself. My father has been my first ever tutor, and my mother has been a tiger-mom. I also thank my extended family, my maternal grandmother, maternal-aunt, and my little brother Anindya. I appreciate my in-laws for believing in me and supporting me in my decisions. I miss the love and warmth of my paternal grandmother on this biggest moment of my life.

Last, but not least, I would like to thank my wife Rituparna Paul for her immense support during the last phase. She left her job and came all the way to be with me so that I do not need to think of anything else and concentrate on completing my dissertation. 

\vskip2cm
\begin{flushleft}
Arpan Mukherjee \\
University at Buffalo \\
January 2018
\end{flushleft}

